Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-09 eurocg/paper.tex
removed some lemmas

\newpage %delete later
\section{Two polynomial special cases}

@@

Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $. If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $. If $ u $ is a block vertex $ a,b $ are the points where the by $ \textit{cycle}(u) $ represented edges intersect $ s(A(p(u))) $. Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}\\
\end{align*}

For an articulation vertex $ u \in \mathcal V $:
\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
\begin{cases}
\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
\sollong(v) & \text{else}
\end{cases}
\end{align}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$. Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle. In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al. \cite{bcgkmn-pped-17}. Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.

Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $. The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $. Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}. Now let $ u $ be a block vertex. We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ p = A(p(u)) $. The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%\begin{lemma}
%	\label{lem:cactus_correctness}	
%	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
%\end{lemma}
%\begin{proof}	
%	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$. Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle. In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al. \cite{bcgkmn-pped-17}. Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
%
%	Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $. The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $. Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}. Now let $ u $ be a block vertex. We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ p = A(p(u)) $. The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%\end{proof}
\begin{lemma}
	\label{lem:compute_block}
	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every vertex $ u \in \mathcal V $ with degree $ \delta $. We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}
\begin{proof}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above. First let $ u \in \mathcal V $ be an articulation vertex, then the result follows from Lemma~\ref{lem:compute_vertex}, since an articulation vertex has at most $ \delta $ many children in $ \mathcal{T} $, one for each cycle it is part of in $ C $ and one for every other articulation vertex it was already connected to in $ C $.

	Now let $ u \in \mathcal V $ be a block vertex. Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n_c) $ time by Bruckdorfer et al. with $ n_c $ being the number of vertices in the cycle. Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ k $ many children, which are all articulation points. Their values are already known and maximization over them can hence be done in $ O(k) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}. Following we need $ O(n_ck) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
\end{proof}

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}. Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block}.
\end{proof}

>>>>

%\newpage %delete later
\section{Two polynomial special cases}

@@

Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $. If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $. If $ u $ is a block vertex $ a,b $ are the points where edges $ s(v), v \in \textit{cycle}(u) $ intersect $ s(A(p(u))) $. Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}
\end{align*}
For an articulation vertex $ u \in \mathcal V $:
\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
\begin{cases}
\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
\sollong(v) & \text{else}
\end{cases}
\end{align}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$. Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle. In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al. \cite{bcgkmn-pped-17}. Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.

Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $. The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $. Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}. Now let $ u $ be a block vertex. We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ p = A(p(u)) $. The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%\begin{lemma}
%	\label{lem:cactus_correctness}	
%	Let $ G $ be a simple graph and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with root $ r \in V $ and maximum degree $ k $, then $ \sollong(r) = \max \{T_i(r) \mid 0 \leq i \leq k \} $ is an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.
%\end{lemma}
%\begin{proof}	
%	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$. Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle. In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al. \cite{bcgkmn-pped-17}. Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
%
%	Now let $ u \in V $ with degree $ \delta $ be an inner vertex of $ \mathcal T $. The values $ \solshort(v) $, $ \solmid(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $. Again first let $ u $ be an articulation vertex, then the maximum for stub $ \ell_i(A(u)) $ is said value plus the sum of the compatible children, which is just Recurrence~\ref{rec:cactus_articulation}. Now let $ u $ be a block vertex. We have to cut open the cycle it represents (compare \cite{bcgkmn-pped-17}). We do this at $ p = A(p(u)) $. The resulting graph is a path over the vertices in $ \textit{cycle}(u) $ with attached articulation vertices. This is especially a tree and we use the aglorithm from Section~\ref{sec:tree}, maximizing over three values of the articulation vertices though,  to compute the $ T_i(u) $.
%\end{proof}
\begin{lemma}
	\label{lem:compute_block}
	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $. We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n_c\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}
\begin{proof}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ with maximum degree $ k $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above. Let $ u \in \mathcal V $ be a block vertex. Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n_c) $ time by Bruckdorfer et al. with $ n_c $ being the number of vertices in the cycle. Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ k $ many children, which are all articulation points. Their values are already known and maximization over them can hence be done in $ O(k) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}. Following we need $ O(n_ck) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
\end{proof}

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Recurrence~\ref{rec:cactus_articulation}. Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_block} for the block vertices and applying Lemma~\ref{lem:compute_vertex} for the articulation vertices.
\end{proof}
